---
title: "04_walkshed_demos_EDA"
author: "Dexter H. Locke, PhD"
date: "`r format(Sys.time())`"
output: html_document
editor_options: 
  chunk_output_type: console
---
calc percent of all in walkshed
calc percent of a race/ethnicity group
then ratio


Numerator is % POC with Access= # POC with access/# POC total
Denominator is % NH with access= # NH with access/# NH white total

then if > 1 better access for BIPOC, if < 1 then worse


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 0 load libraries, get oriented
```{r}

# Load libraries 
packs <-c(  'tidyverse' # cuz
          , 'tidylog'   # prints out what was done in dplyr and tidyr; VERBOSE
          , 'magrittr'  # for all of the the pipes
          , 'sf'        # for spatial data support
          # , 'sp'        #  older spatial data support
          # , 'spdep'     # spatial dependency
          , 'sfdep'     # remotes::install_github("josiahparry/sfdep")
          , 'mapview'   # web maps for zooming and panning around
          , 'beepr'     # makes noises
          , 'tictoc'    # times things
          , 'tigris'    # Census geographic data (via TIGER)
          , 'parallel'  # parallel processing, vroom, vroom
          )

if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packs, rownames(installed.packages())))
}

# load the packages all at once
vapply(packs, library, character.only = TRUE, logical(1),
       logical.return = TRUE, quietly = TRUE)

# get oriented
list.files()
list.files('input_data')


# for reproducibility
set.seed(19870630)

# fixes mapview
mapviewOptions(fgb = FALSE)
# mapviewOptions(platform = "leafgl") #should help with large data.

# custom function for "Not In"
`%nin%` <- Negate(`%in%`)


# function for reclassifying Census codes to graph/table-friendly labels
handsome_cats <- function(data, var){
  mutate(
    data
    , {{ var }} := str_replace_all({{ var }}, 'Hisp'          , 'Hispanic / Latinx')
    , {{ var }} := str_replace_all({{ var }}, 'NH_AIAN_alone' , 'NH-Native American and Alaska Native alone')
    , {{ var }} := str_replace_all({{ var }}, 'NH_Asian_alone', 'NH-Asian alone')
    , {{ var }} := str_replace_all({{ var }}, 'NH_Black_alone', 'NH-Black or African American alone')
    , {{ var }} := str_replace_all({{ var }}, 'NH_multirace'  , 'NH two or more races')
    , {{ var }} := str_replace_all({{ var }}, 'NH_NHPI_alone' , 'NH-Native Hawaiian and Pacific Islander')
    , {{ var }} := str_replace_all({{ var }}, 'NH_other'      , 'NH-Other')
    , {{ var }} := str_replace_all({{ var }}, 'NH_White_alone', 'NH-White alone')
    , {{ var }} := str_replace_all({{ var }}, 'tot_pop'       , 'Total Population')
  )
}


```



# 1 read in data
## A block-walkshed intersection table from meta-batch pieces
```{r}

# get file names
(files     <- dir(  'input_data/walkshed_block_intersections'
                 , recursive = TRUE, pattern = "*.csv$", full.names = TRUE))

removals <- c('input_data/walkshed_block_intersections/walkshed_block_intersections_',
              '_1_to_43.csv', '_44_to_86.csv', '_87_to_130.csv', '_131_to_172.csv')


tic(); (walkshed_block_int_tbl <- tibble(filename = files) %>%
  mutate(
    file_contents = map(files, 
                        ~read_csv(.
                                  , progress = TRUE
                                  , col_names = c('sesync_id', 'GISJOIN', 'prop_in')
                                  ))) |> 
  unnest(cols = file_contents) |> 
  # needed?
  # mutate(meta_batch = as.factor(str_remove_all(filename, paste(removals, collapse = '|')))) |> 
  select(-filename)); toc()

# 2x check proportion behaves - it does!
walkshed_block_int_tbl |> summary(prop_in)
walkshed_block_int_tbl |> 
  ggplot(aes(prop_in)) +
  geom_density()

(problem_ints <- walkshed_block_int_tbl |> filter(prop_in > 1) |> arrange(desc(prop_in)))

# problem_ints |> head(20) |> View()

```



## B Census blocks: tabular 
```{r}

tic(); (block_tbl <- read_csv('input_data/nhgis0059_csv/nhgis0059_ds248_2020_block.csv'
                    , col_select = c(  GISJOIN
                                     , year      = YEAR
                                     , state     = STATE
                                     , state_id  = STATEA
                                     , county    = COUNTY
                                     , area_land = AREALAND
                                     , area_water= AREAWATR
                                     , tot_pop   = U7C001     # Total population
                                     , Hisp      = U7C002     # Hispanic or Latino
                                     # NH White alone (not POC). Create a %nhwhite variable
                                     # that is U7C005/U7C001. This may be prime variable of
                                     # interest.
                                     , NH_White_alone = U7C005
                                     , NH_Black_alone = U7C006      
                                     , NH_AIAN_alone  = U7C007      
                                     , NH_Asian_alone = U7C008      
                                     , NH_NHPI_alone  = U7C009      
                                     , NH_other       = U7C010      
                                     , NH_multirace   = U7C011      
                                     )
                    )); toc()

```



## C block-Urban Area look up
```{r}


(ua_assignment_path <- paste0(getwd(), '/input_data/combined_blocks_2020_no_water_single_part_Urban_Area_lookup/block_to_ua_lookup.csv'))

(ua_assignment <- read_csv(ua_assignment_path
                           , col_names = c('mult_id', 'GISJOIN', 'UACE10', 'UA_name')))

```



## D Census divisions / regions polygons
```{r}

(
  div_regions <- readxl::read_xlsx('input_data/state-geocodes-v2017.xlsx', skip = 4) |> 
    mutate(div_name = ifelse(Division == '0', Name, NA),
           region_name = ifelse(Division != '0' & `State (FIPS)` == '00', Name, NA)) |> 
    fill(div_name, region_name) |> 
    filter(`State (FIPS)` != '00')
 )

# 2x check
div_regions |> distinct(div_name, region_name)
div_regions |> distinct(Region, Division)

```



# 2 join tabular blocks, intersection file, urban areas, divisions/ regions
```{r eval=FALSE, include=FALSE}
# # in-memory tests
# walkshed_block_int_tbl |> left_join(block_tbl, by = 'GISJOIN')
# walkshed_block_int_tbl |> anti_join(block_tbl, by = 'GISJOIN')

tic(); (
  blocks_tbl_augmented <- block_tbl |>
    left_join(walkshed_block_int_tbl, by = 'GISJOIN') |>                    # walk Ints +sesync_ids
    left_join(ua_assignment |> select(GISJOIN, UA_name), by = 'GISJOIN') |> # Urban Areas
    left_join(div_regions |>                                                # divs / regions
              select(state_id = `State (FIPS)`, div_name, region_name), by = 'state_id')); toc() # ~20s
```



# 3 Demo summaries 
## A data-prep: whole country and extract non-walkshed blocks
```{r}
# # blocks_tbl_augmented |> 
# non_ws_demos <- block_tbl |> 
#   # filter(state == 'Rhode Island') |> # just used for testing
#   left_join(walkshed_block_int_tbl, by = 'GISJOIN') |> 
#   filter(is.na(sesync_id)) #|> # finds blocks not in walkshed

# get whole-country demographics
(
  USA_total_demos <- block_tbl |> 
    select(GISJOIN, tot_pop : NH_multirace) |> 
    pivot_longer(-GISJOIN) |> 
    group_by(name) |> 
    # summarise(n = sum(value, na.rm = TRUE), source = 'usa_total_demos')
    summarise(usa_total_demos = sum(value, na.rm = TRUE))
  )

# test out custom labeling function.
USA_total_demos |> handsome_cats(name)
# USA_total_demos |> handsome_cats(name) |>
#   write_csv(paste0(getwd(), '/output_tables/USA_total_demos_', Sys.Date(), '.csv'))


# finds blocks NOT in walkshed
(
  non_ws_demos <- block_tbl |> 
    # filter(state == 'Rhode Island') |> # just used for testing
    anti_join(walkshed_block_int_tbl, by = 'GISJOIN')
  )


# 2x checks
non_ws_demos |> glimpse()
non_ws_demos |> summary()
non_ws_demos |> map(~sum(is.na(.))) |> bind_rows() |> t()

```


## B national access
```{r}

# get non-walkshed demographics for the country
(
  non_ws_demos_sum <- non_ws_demos |> 
    select(GISJOIN, tot_pop : NH_multirace) |> 
    pivot_longer(-GISJOIN) |> 
    group_by(name) |> 
    # summarise(n = sum(value, na.rm = TRUE), source = 'non_ws_demos')
    summarise(non_ws_demos = sum(value, na.rm = TRUE))
  )


# combine non-walkshed and whole country in wide format, calc ws demos as difference btw non and whole
(
  combined_demos_wide <- non_ws_demos_sum |>
    left_join(USA_total_demos) |>
    mutate(ws_demos = usa_total_demos - non_ws_demos
           # , ratio = ws_demos / non_ws_demos
           )
  )

# # save out
# combined_demos_wide |> 
#   handsome_cats(name) |>
#   write_csv(paste0(getwd(), '/output_tables/combined_demos_wide_', Sys.Date(), '.csv'))


(combined_demos_wide |> 
  # select(name, ws_demos) |> 
  filter(name != 'tot_pop') |> 
  handsome_cats(name) |> 
  mutate(cat = ifelse(name == 'NH-White alone', 'NH-White alone', 'BIPOC')) |> 
  group_by(cat) |> 
  summarise(         sum_ws_demos = sum(ws_demos)
            , sum_usa_total_demos = sum(usa_total_demos)) |> 
  mutate(frac = sum_ws_demos / sum_usa_total_demos) |> 
  select(cat, frac) |> 
  t() |> as_tibble() |> 
  janitor::row_to_names(row_number = 1) |> 
  mutate_all(as.numeric) |> 
  mutate(ratio = BIPOC / `NH-White alone`) -> has_access_national_ratio)
  
# has_access_national_ratio |> 
#   write_csv(paste0(getwd(), '/output_tables/has_access_national_ratio_', Sys.Date(), '.csv'))


# make long form for graphing
(
  combined_demos_long <- combined_demos_wide |> 
    select(name, non_ws_demos, ws_demos) |> 
    pivot_longer(-name, names_to = 'source')
  )

# # save out
# combined_demos_long |>
#   handsome_cats(name) |>
#   write_csv(paste0(getwd(), '/output_tables/combined_demos_long_', Sys.Date(), '.csv'))


# viz
combined_demos_long |> 
  mutate(' ' = fct_recode(factor(source)
                          , `Oustside 10-minute walk` = 'non_ws_demos'
                          , `Within 10-minute walk` = 'ws_demos'
                          )) |> 
  # filter(name != 'tot_pop') |> 
  handsome_cats(name) |> 
  ggplot(aes(value, name, fill = ` `)) + 
  geom_col(position = 'dodge') + 
  theme_bw(18) + 
  labs(  title = 'Park Access by Race/Ethnicity'
       , subtitle = 'entire USA'
       , y = ''
       , x = 'Estimated number of people'
       , caption = 'NH = Non-Hispanic\nData Sources: Trust for Public Land walkshed;\nU.S. Census Bureau 2020 blocks') +
  scale_x_continuous(labels = scales::comma, expand = c(0, 0), lim = c(0, 2e8)) +
  scale_fill_viridis_d() + 
  # theme(legend.position = 'bottom') + 
  NULL

# # save it out
# ggsave(file = paste0(getwd(), '/figures/usa_park_access_10min_', Sys.Date(), '.png')
#        , width = 7*2, height = 4*2)


# people WITHOUT ACCESS
non_ws_demos_sum |> 
  filter(name != 'tot_pop') |> 
  handsome_cats(name) |> 
  ggplot(aes(non_ws_demos, name)) +
  geom_col() + 
  theme_bw(16) + 
  scale_x_continuous(labels = scales::comma) +
  NULL


```


## C urban vs not
```{r}

# get whole-country demographics
(
  USA_total_demos_urban_or_not_summ <- block_tbl |> 
    left_join(ua_assignment |> select(GISJOIN, UA_name), by = 'GISJOIN') |> # Urban Areas
    mutate(urban = ifelse(is.na(UA_name), 'not urban', 'urban')) |> 
    select(GISJOIN, urban, tot_pop : NH_multirace) |> 
    pivot_longer(cols = c(tot_pop : NH_multirace)) |>
    group_by(urban, name) |> 
    # summarise(n = sum(value, na.rm = TRUE), source = 'usa_total_demos')
    summarise(usa_total_demos = sum(value, na.rm = TRUE))
  )


( # NOT IN WALKSHEDS
  non_ws_demos_urban_summ <- block_tbl |> 
  # filter(state == 'Rhode Island') |> # just used for testing
    anti_join(walkshed_block_int_tbl, by = 'GISJOIN') |>                    # finds blocks NOT in walkshed
    left_join(ua_assignment |> select(GISJOIN, UA_name), by = 'GISJOIN') |> # Urban Areas
    mutate(urban = ifelse(is.na(UA_name), 'not urban', 'urban')) |> 
    select(GISJOIN, urban, tot_pop : NH_multirace) |> 
    pivot_longer(cols = c(tot_pop : NH_multirace)) |>
    group_by(urban, name) |> 
    summarise(non_ws_demos = sum(value, na.rm = TRUE))
  )

# combine non-walkshed and whole country in wide format, calc ws demos as difference btw non and whole
(
  combined_demos_urban_wide <- non_ws_demos_urban_summ |>
    left_join(USA_total_demos_urban_or_not_summ, by = c('urban', 'name')) |>
    mutate(ws_demos = usa_total_demos - non_ws_demos
           # , ratio = ws_demos / non_ws_demos
           ) |> 
    relocate(usa_total_demos, .after = last_col())
  )

# # write out
# combined_demos_urban_wide |>
#   handsome_cats(name) |>
#   write_csv(paste0(getwd(), '/output_tables/combined_demos_urban_wide_', Sys.Date(), '.csv'))


(
  combined_demos_urban_wide |> 
    filter(name != 'tot_pop') |> 
    handsome_cats(name) |> 
    mutate(cat = ifelse(name == 'NH-White alone', 'NH-White alone', 'BIPOC')) |> 
    group_by(cat, urban) |>
    summarise(         sum_ws_demos = sum(ws_demos)
              , sum_usa_total_demos = sum(usa_total_demos)) |>
    mutate(frac = sum_ws_demos / sum_usa_total_demos) |> 
    select(cat, urban, frac) |> 
    pivot_wider(id_cols = urban, names_from = cat, values_from = frac) |> 
    mutate(ratio = BIPOC / `NH-White alone`) -> has_access_urban_ratio
  )

# # write out
# has_access_urban_ratio |>
#   write_csv(paste0(getwd(), '/output_tables/has_access_urban_ratio_', Sys.Date(), '.csv'))



# make long form for graphing
(
  combined_demos_urban_long <- combined_demos_urban_wide |> 
    select(name, urban, non_ws_demos, ws_demos) |> 
    pivot_longer(cols = c(non_ws_demos, ws_demos), names_to = 'source')
  )

# # write out
# combined_demos_urban_long |> 
#   handsome_cats(name) |>
#   write_csv(paste0(getwd(), '/output_tables/combined_demos_urban_long_', Sys.Date(), '.csv'))


# viz urban vs not, within vs outside walk by race/ethnicity
combined_demos_urban_long |> 
  mutate(' ' = fct_recode(factor(source)
                          , `Oustside 10-minute walk` = 'non_ws_demos'
                          , `Within 10-minute walk` = 'ws_demos'
                          )) |> 
  # filter(name != 'tot_pop') |> 
  handsome_cats(name) |> 
  ggplot(aes(value, name, fill = ` `)) + 
  geom_col(position = 'dodge') + 
  theme_bw(18) + 
  labs(  title = 'Park Access by Race/Ethnicity and Urbanicity'
       # , subtitle = ''
       , y = ''
       , x = 'Estimated number of people'
       , caption = 'NH = Non-Hispanic\nData Sources: Trust for Public Land walkshed;\nU.S. Census Bureau 2020 blocks') +
  # scale_x_continuous(labels = scales::comma, expand = c(0, 0), lim = c(0, 2e8)) +
  scale_x_continuous(labels = scales::label_number(suffix = " M", scale = 1e-6)
                     # , expand = c(0, 0)
                     ) + # millions
  scale_fill_viridis_d() + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5)) +
  facet_wrap(~urban) + 
  theme(legend.position = 'bottom') + 
  NULL

# # save it out
# ggsave(file = paste0(getwd(), '/figures/USA_park_access_10min_urban_vs_not_', Sys.Date(), '.png')
#        , width = 7*2, height = 4*2)


```


### i urban only
```{r}

# viz urban only, within vs outside walk by race/ethnicity
combined_demos_urban_long |> 
  mutate(' ' = fct_recode(factor(source)
                          , `Oustside 10-minute walk` = 'non_ws_demos'
                          , `Within 10-minute walk` = 'ws_demos'
                          )) |> 
  # filter(name != 'tot_pop') |> 
  handsome_cats(name) |> 
  filter(urban == 'urban') |> 
  ggplot(aes(value, name, fill = ` `)) + 
  geom_col(position = 'dodge') + 
  theme_bw(18) + 
  labs(  title = 'Park Access by Race/Ethnicity'
       , subtitle = 'Urban Areas only'
       , y = ''
       , x = 'Estimated number of people'
       , caption = 'NH = Non-Hispanic\nData Sources: Trust for Public Land walkshed;\nU.S. Census Bureau 2020 blocks') +
  # scale_x_continuous(labels = scales::comma, expand = c(0, 0), lim = c(0, 2e8)) +
  scale_x_continuous(labels = scales::label_number(suffix = " M", scale = 1e-6)
                     , expand = c(0, 0)
                     , limits = c(0, 2e8)
                     ) + # millions
  scale_fill_viridis_d() + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5)) +
  theme(legend.position = 'bottom') + 
  NULL

# # save it out
# ggsave(file = paste0(getwd(), '/figures/usa_park_access_10min_urban_only_', Sys.Date(), '.png')
#        , width = 7*2, height = 4*2)

```


## D Divisions - urban only
```{r}

# get whole-country demographics by urban and division
(
  USA_total_demos_divisions_urban_summ <- block_tbl |> 
    left_join(ua_assignment |> select(GISJOIN, UA_name), by = 'GISJOIN') |> # Urban Areas
    mutate(urban = ifelse(is.na(UA_name), 'not urban', 'urban')) |> 
    left_join(div_regions |>                                                # divisions
              select(state_id = `State (FIPS)`, Division = div_name), by = 'state_id') |> 
    select(GISJOIN, Division, urban, tot_pop : NH_multirace) |> 
    pivot_longer(cols = c(tot_pop : NH_multirace)) |>
    group_by(Division, urban, name) |> 
    summarise(usa_total_demos = sum(value, na.rm = TRUE))
  )


( # NOT IN WALKSHEDS
  non_ws_demos_divisions_urban_summ <- block_tbl |> 
  # filter(state == 'Rhode Island') |> # just used for testing
    anti_join(walkshed_block_int_tbl, by = 'GISJOIN') |>              # finds blocks NOT in walkshed
    left_join(ua_assignment |> select(GISJOIN, UA_name), by = 'GISJOIN') |> # Urban Areas
    mutate(urban = ifelse(is.na(UA_name), 'not urban', 'urban')) |> 
    left_join(div_regions |>                                                # divisions
                select(state_id = `State (FIPS)`, Division = div_name), by = 'state_id') |> 
    select(GISJOIN, Division, urban, tot_pop : NH_multirace) |> 
    pivot_longer(cols = c(tot_pop : NH_multirace)) |>
    group_by(Division, urban, name) |> 
    summarise(non_ws_demos = sum(value, na.rm = TRUE))
  )

# combine non-walkshed and whole country in wide format, calc ws demos as difference btw non and whole
(
  combined_demos_divisions_urban_wide <- non_ws_demos_divisions_urban_summ |>
    left_join(USA_total_demos_divisions_urban_summ, by = c('Division', 'urban', 'name')) |>
    mutate(ws_demos = usa_total_demos - non_ws_demos
           # , ratio = ws_demos / non_ws_demos
           ) |> 
    relocate(usa_total_demos, .after = last_col())
  )

# # write out
# combined_demos_divisions_urban_wide |>
#   handsome_cats(name) |>
#   write_csv(paste0(getwd(), '/output_tables/combined_demos_divisions_urban_wide_', Sys.Date(), '.csv'))


(
  combined_demos_divisions_urban_wide |>
    filter(name != 'tot_pop') |> 
    filter(!is.na(Division)) |>
    handsome_cats(name) |> 
    mutate(cat = ifelse(name == 'NH-White alone', 'NH-White alone', 'BIPOC')) |> 
    group_by(cat, urban, Division) |>
    summarise(         sum_ws_demos = sum(ws_demos)
              , sum_usa_total_demos = sum(usa_total_demos)) |>
    mutate(frac = sum_ws_demos / sum_usa_total_demos) |> 
    select(cat, urban, Division, frac) |> 
    pivot_wider(id_cols = c(urban, Division), names_from = cat, values_from = frac) |> 
    mutate(ratio = BIPOC / `NH-White alone`) -> has_access_urban_divisions_ratio
  )

# has_access_urban_divisions_ratio |> 
#   write_csv(paste0(getwd(), '/output_tables/has_access_urban_divisions_ratio_', Sys.Date(), '.csv'))



has_access_urban_divisions_ratio |> 
  ggplot(aes(ratio, Division)) +
  geom_col() + 
  theme_bw(18) +
  labs(  title = 'Park Access by Race/Ethnicity and Division'
       # , subtitle = 'Urban areas only'
       , y = ''
       , x = 'Ratio of BIPOC with park access to NH-White with parck access'
       , caption = 'NH = Non-Hispanic\nData Sources: Trust for Public Land walkshed;\nU.S. Census Bureau 2020 blocks') +
  scale_x_continuous(expand = c(0, 0), lim = c(0, 1.45)) +
  # scale_x_continuous(labels = scales::label_number(suffix = " M", scale = 1e-6)
  #                    , expand = c(0, 0), limits = c(0, 6e7)
  #                    ) + # millions
  scale_fill_viridis_d() +
  # theme(axis.text.x = element_text(angle = 90, vjust = .5)) +
  # theme(legend.position = 'bottom') + 
  geom_vline(xintercept = 1) + 
  facet_wrap(~urban) + 
  NULL

# # save it out
# ggsave(file = paste0(getwd(), '/figures/Divisions_park_access_10min_ratio_by_urban_'
#                      , Sys.Date(), '.png')
#        , width = 7*2, height = 4*2)




# make long form for graphing
(
  combined_demos_divisions_urban_long <- combined_demos_divisions_urban_wide |> 
    select(name, urban, non_ws_demos, ws_demos) |> 
    pivot_longer(cols = c(non_ws_demos, ws_demos), names_to = 'source')
  )

# # write out
# combined_demos_divisions_urban_long |> 
#   handsome_cats(name) |>
#   write_csv(paste0(getwd(), '/output_tables/combined_demos_divisions_urban_long_', Sys.Date(), '.csv'))


# viz urban vs not, within vs outside walk by race/ethnicity
combined_demos_divisions_urban_long |> 
  mutate(' ' = fct_recode(factor(source)
                          , `Oustside 10-minute walk` = 'non_ws_demos'
                          , `Within 10-minute walk` = 'ws_demos'
                          )) |> 
  # filter(name != 'tot_pop') |> 
  handsome_cats(name) |> 
  filter(urban == 'urban') |> 
  ggplot(aes(value, name, fill = ` `)) + 
  geom_col(position = 'dodge') + 
  theme_bw(18) + 
  labs(  title = 'Park Access by Race/Ethnicity and Division'
       , subtitle = 'Urban areas only'
       , y = ''
       , x = 'Estimated number of people'
       , caption = 'NH = Non-Hispanic\nData Sources: Trust for Public Land walkshed;\nU.S. Census Bureau 2020 blocks') +
  # scale_x_continuous(labels = scales::comma, expand = c(0, 0), lim = c(0, 2e8)) +
  scale_x_continuous(labels = scales::label_number(suffix = " M", scale = 1e-6)
                     , expand = c(0, 0), limits = c(0, 6e7)
                     ) + # millions
  scale_fill_viridis_d() + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5)) +
  theme(legend.position = 'bottom') + 
  facet_wrap(~Division) + 
  NULL

# # save it out
# ggsave(file = paste0(getwd(), '/figures/Divisions_park_access_10min_urban_only_', Sys.Date(), '.png')
#        , width = 7*2, height = 4*2)


```



## E Regions - urban only
```{r}

# get whole-country demographics by urban and region
(
  USA_total_demos_regions_urban_summ <- block_tbl |> 
    left_join(ua_assignment |> select(GISJOIN, UA_name), by = 'GISJOIN') |> # Urban Areas
    mutate(urban = ifelse(is.na(UA_name), 'not urban', 'urban')) |> 
    left_join(div_regions |>                                                # divisions
              select(state_id = `State (FIPS)`, Region = region_name), by = 'state_id') |> 
    select(GISJOIN, Region, urban, tot_pop : NH_multirace) |> 
    pivot_longer(cols = c(tot_pop : NH_multirace)) |>
    group_by(Region, urban, name) |> 
    summarise(usa_total_demos = sum(value, na.rm = TRUE))
  )


( # NOT IN WALKSHEDS
  non_ws_demos_regions_urban_summ <- block_tbl |> 
  # filter(state == 'Rhode Island') |> # just used for testing
    anti_join(walkshed_block_int_tbl, by = 'GISJOIN') |>              # finds blocks NOT in walkshed
    left_join(ua_assignment |> select(GISJOIN, UA_name), by = 'GISJOIN') |> # Urban Areas
    mutate(urban = ifelse(is.na(UA_name), 'not urban', 'urban')) |> 
    left_join(div_regions |>                                                # divisions
              select(state_id = `State (FIPS)`, Region = region_name), by = 'state_id') |> 
    select(GISJOIN, Region, urban, tot_pop : NH_multirace) |> 
    pivot_longer(cols = c(tot_pop : NH_multirace)) |>
    group_by(Region, urban, name) |> 
    summarise(non_ws_demos = sum(value, na.rm = TRUE))
  )

# combine non-walkshed and whole country in wide format, calc ws demos as difference btw non and whole
(
  combined_demos_regions_urban_wide <- non_ws_demos_regions_urban_summ |>
    left_join(USA_total_demos_regions_urban_summ, by = c('Region', 'urban', 'name')) |>
    mutate(ws_demos = usa_total_demos - non_ws_demos
           # , ratio = ws_demos / non_ws_demos
           ) |> 
    relocate(usa_total_demos, .after = last_col())
  )

# # write out
# combined_demos_regions_urban_wide |>
#   handsome_cats(name) |>
#   write_csv(paste0(getwd(), '/output_tables/combined_demos_regions_urban_wide_', Sys.Date(), '.csv'))



(
  combined_demos_regions_urban_wide |>
    filter(name != 'tot_pop') |> 
    filter(!is.na(Region)) |>
    handsome_cats(name) |> 
    mutate(cat = ifelse(name == 'NH-White alone', 'NH-White alone', 'BIPOC')) |> 
    group_by(cat, urban, Region) |>
    summarise(         sum_ws_demos = sum(ws_demos)
              , sum_usa_total_demos = sum(usa_total_demos)) |>
    mutate(frac = sum_ws_demos / sum_usa_total_demos) |> 
    select(cat, urban, Region, frac) |> 
    pivot_wider(id_cols = c(urban, Region), names_from = cat, values_from = frac) |> 
    mutate(ratio = BIPOC / `NH-White alone`) -> has_access_urban_regions_ratio
  )

# # write out
# has_access_urban_regions_ratio |>
#   write_csv(paste0(getwd(), '/output_tables/has_access_urban_regions_ratio_', Sys.Date(), '.csv'))



has_access_urban_regions_ratio |> 
  ggplot(aes(ratio, Region)) +
  geom_col() + 
  theme_bw(18) +
  labs(  title = 'Park Access by Race/Ethnicity and Region'
       # , subtitle = 'Urban areas only'
       , y = ''
       , x = 'Ratio of BIPOC with park access to NH-White with parck access'
       , caption = 'NH = Non-Hispanic\nData Sources: Trust for Public Land walkshed;\nU.S. Census Bureau 2020 blocks') +
  scale_x_continuous(expand = c(0, 0), lim = c(0, 1.75)) +
  # scale_x_continuous(labels = scales::label_number(suffix = " M", scale = 1e-6)
  #                    , expand = c(0, 0), limits = c(0, 6e7)
  #                    ) + # millions
  scale_fill_viridis_d() +
  # theme(axis.text.x = element_text(angle = 90, vjust = .5)) +
  # theme(legend.position = 'bottom') + 
  geom_vline(xintercept = 1) + 
  facet_wrap(~urban) + 
  NULL

# # save it out
# ggsave(file = paste0(getwd(), '/figures/Regions_park_access_10min_ratio_by_urban_'
#                      , Sys.Date(), '.png')
#        , width = 7*2, height = 4*2)





# make long form for graphing
(
  combined_demos_regions_urban_long <- combined_demos_regions_urban_wide |> 
    select(name, urban, non_ws_demos, ws_demos) |> 
    pivot_longer(cols = c(non_ws_demos, ws_demos), names_to = 'source')
  )

# # write out
# combined_demos_regions_urban_long |> 
#   handsome_cats(name) |>
#   write_csv(paste0(getwd(), '/output_tables/combined_demos_regions_urban_long_', Sys.Date(), '.csv'))


# viz urban vs not, within vs outside walk by race/ethnicity
combined_demos_regions_urban_long |> 
  mutate(' ' = fct_recode(factor(source)
                          , `Oustside 10-minute walk` = 'non_ws_demos'
                          , `Within 10-minute walk` = 'ws_demos'
                          )) |> 
  # filter(name != 'tot_pop') |> 
  handsome_cats(name) |> 
  filter(urban == 'urban') |> 
  ggplot(aes(value, name, fill = ` `)) + 
  geom_col(position = 'dodge') + 
  theme_bw(18) + 
  labs(  title = 'Park Access by Race/Ethnicity and Region'
       , subtitle = 'Urban areas only'
       , y = ''
       , x = 'Estimated number of people'
       , caption = 'NH = Non-Hispanic\nData Sources: Trust for Public Land walkshed;\nU.S. Census Bureau 2020 blocks') +
  # scale_x_continuous(labels = scales::comma, expand = c(0, 0), lim = c(0, 2e8)) +
  scale_x_continuous(labels = scales::label_number(suffix = " M", scale = 1e-6)
                     , expand = c(0, 0), limits = c(0, 45e6)
                     ) + # millions
  scale_fill_viridis_d() + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5)) +
  theme(legend.position = 'bottom') + 
  facet_wrap(~str_wrap(Region, 20)) + 
  NULL

# # save it out
# ggsave(file = paste0(getwd(), '/figures/Regions_park_access_10min_urban_only_', Sys.Date(), '.png')
#        , width = 7*2, height = 4*3 )


```



## save out

```{r eval=FALSE, include=FALSE}
# save.image("saved_sessions/04_walkshed_demos_EDA.RData") 
```

## CITE your sources!!!

```{r}
lapply(packs, citation); rm(packs)
sessionInfo()
```

Last Knit on `r format(Sys.time())`

# SANDBOX
```{r}


# mutate(  p_tot_pop        = ifelse(tot_pop == 0, 0, 100*(tot_pop        / tot_pop))
  #        , p_Hisp           = ifelse(tot_pop == 0, 0, 100*(Hisp           / tot_pop))
  #        , p_NH_White_alone = ifelse(tot_pop == 0, 0, 100*(NH_White_alone / tot_pop))
  #        , p_NH_Black_alone = ifelse(tot_pop == 0, 0, 100*(NH_Black_alone / tot_pop))
  #        , p_NH_AIAN_alone  = ifelse(tot_pop == 0, 0, 100*(NH_AIAN_alone  / tot_pop))
  #        , p_NH_Asian_alone = ifelse(tot_pop == 0, 0, 100*(NH_Asian_alone / tot_pop))
  #        , p_NH_NHPI_alone  = ifelse(tot_pop == 0, 0, 100*(NH_NHPI_alone  / tot_pop))
  #        , p_NH_other       = ifelse(tot_pop == 0, 0, 100*(NH_other       / tot_pop))
  #        , p_NH_multirace   = ifelse(tot_pop == 0, 0, 100*(NH_multirace   / tot_pop))
  #        , sum_checker = p_Hisp + p_NH_White_alone + p_NH_Black_alone + p_NH_AIAN_alone + p_NH_Asian_alone + p_NH_NHPI_alone + p_NH_other + p_NH_multirace 
  #        ) 

```





# ?make walkshed-level estimates
```{r}
# make walkshed-level estimates
block_tbl |> 
  right_join(walkshed_block_int_tbl, by = 'GISJOIN') |> # filter(is.na(sesync_id))
  mutate(  w_tot_pop        = tot_pop       *prop_in
         , w_Hisp           = Hisp          *prop_in
         , w_NH_White_alone = NH_White_alone*prop_in
         , w_NH_Black_alone = NH_Black_alone*prop_in     
         , w_NH_AIAN_alone  = NH_AIAN_alone *prop_in      
         , w_NH_Asian_alone = NH_Asian_alone*prop_in     
         , w_NH_NHPI_alone  = NH_NHPI_alone *prop_in      
         , w_NH_other       = NH_other      *prop_in      
         , w_NH_multirace   = NH_multirace  *prop_in
         ) |> 
  # arrange(GISJOIN, sesync_id) |> # cosmetic, helps with debugging
  group_by(sesync_id) |> 
  summarise(  est_tot_pop        = sum(w_tot_pop)
            , est_Hisp           = sum(w_Hisp)
            , est_NH_White_alone = sum(w_NH_White_alone)
            , est_NH_Black_alone = sum(w_NH_Black_alone)
            , est_NH_AIAN_alone  = sum(w_NH_AIAN_alone)
            , est_NH_Asian_alone = sum(w_NH_Asian_alone)
            , est_NH_NHPI_alone  = sum(w_NH_NHPI_alone)
            , est_NH_other       = sum(w_NH_other)
            , est_NH_multirace   = sum(w_NH_multirace)
            ) |> 
  mutate(  p_tot_pop        = 100*(est_tot_pop        / est_tot_pop)
         , p_Hisp           = 100*(est_Hisp           / est_tot_pop)
         , p_NH_White_alone = 100*(est_NH_White_alone / est_tot_pop)
         , p_NH_Black_alone = 100*(est_NH_Black_alone / est_tot_pop)
         , p_NH_AIAN_alone  = 100*(est_NH_AIAN_alone  / est_tot_pop)
         , p_NH_Asian_alone = 100*(est_NH_Asian_alone / est_tot_pop)
         , p_NH_NHPI_alone  = 100*(est_NH_NHPI_alone  / est_tot_pop)
         , p_NH_other       = 100*(est_NH_other       / est_tot_pop)
         , p_NH_multirace   = 100*(est_NH_multirace   / est_tot_pop)
         , sum_checker = p_Hisp + p_NH_White_alone + p_NH_Black_alone + p_NH_AIAN_alone + p_NH_Asian_alone + p_NH_NHPI_alone + p_NH_other + p_NH_multirace 
         ) -> ws_demos


ws_demos |> glimpse()

ws_demos |> summary()
ws_demos |> map(~sum(is.na(.))) |> bind_rows() |> t()
ws_demos |> filter(is.na(p_tot_pop)) |> pull(sesync_id) -> walksheds_no_demos

ws |> 
  filter(sesync_id %in% walksheds_no_demos) |> 
  mapview()


```


## ?C how do demos compare?
```{r}

(
  ws_demos |> 
    select(sesync_id, starts_with('est_')) |> 
    pivot_longer(-sesync_id) |> 
    group_by(name) |> 
    summarise(n = sum(value), source = 'TPL walksheds') |> 
    mutate(name = str_remove_all(name, 'est_')) -> ws_demos_sum
  )

ws_demos_sum |> 
  filter(name != 'tot_pop') |> 
  ggplot(aes(n, name)) +
  geom_bar(stat = 'identity') + 
  theme_bw(16) + 
  scale_x_continuous(labels = scales::comma) +
  NULL


(
  block_tbl |>
    select(GISJOIN, tot_pop: NH_multirace) |> 
    pivot_longer(-GISJOIN) |> 
    # filter(name != 'tot_pop') |> 
    group_by(name) |> 
    summarise(n = sum(value), source = 'Whole Country') -> us_demos_sum
  )


bind_rows(ws_demos_sum, us_demos_sum) |> 
  filter(name != 'tot_pop') |> 
  ggplot(aes(n, name, fill = source)) +
  geom_bar(stat = 'identity', position = 'dodge') + 
  theme_bw(16) + 
  scale_x_continuous(labels = scales::comma) +
  NULL

```




